---
title: "Lab Firebrand Module 2"
output: html_notebook
---

First, import the three datasets downloaded from kaggle;
IBM HR Analytics Employee Attrition & Performance
Stack Overflow Developer Survey, 2017
E-Commerce Data
```{r}
#good tools 
devtools::install_github("krupanss/IEDA") # an interactive IDE to explore data
library(IEDA)
runIEDA()
install.packages(c("xray","skimr","visdat","GGally"))
skimr::skim(StackOverflowSurvey)

```
#Exploration and manipulation of IBM HR Survey
```{r}
runIEDA()
skimr::skim(IBMSurvey)
xray::anomalies(IBMSurvey)
```
#Exploration and manipulation of Stack Overflow Developer Survey 2017

Common things to explore/check for:
Unusual distributions
Outliers
Missings
Correlations
```{r}

library(assertive)
library()
library(DataExplorer)
runIEDA() # good tool for understanding the data
skimr::skim(StackOverflowSurvey) #
xray::anomalies(StackOverflowSurvey) #show some variables that probably don't fit in the model
# ExCoder10Years, ExCoderNotForMe, ExCoderWillNotCode, ExCoderReturn, ExCoderBalance, ExCoderBelonged, ExCoderActive, ExCoderSkills, YearsCodedJobPast, MobileDeveloperType, ExpectedSalary, TimeAfterBootcamp, NonDeveloper
xray::distributions(StackOverflowSurvey)

StackOverflowSurvey %>% select(-ExCoder10Years,-ExCoderNotForMe, -ExCoderWillNotCode, -ExCoderReturn, -ExCoderBalance, -ExCoderBelonged, -ExCoderActive, -ExCoderSkills, -YearsCodedJobPast, -MobileDeveloperType, -ExpectedSalary, -TimeAfterBootcamp, -NonDeveloperType)

DataExplorer::creat


```
#Exploration and manipulation of Ecommerce
```{r}
library(tidyverse)
runIEDA()

#"ecommerce" %>%
#  read.csv(progress=FALSE)%>%
 Ecommerce %>%
   mutate(Description=str_remove_all(Description,"[^[:alnum:]]")) %>%
   skimr::skim()
 
skimr::skim(Ecommerce)
xray::anomalies(Ecommerce)
```
