---
title: "R Basics Firebrand"
output:
  html_notebook: default
---

# About R
- R came from S in the 70s
- R works in memory
    + Microsoft has out-of-memory processing capabilities
    + There's some other potential potential options
- R is querky
- R is case sensitive
- R can be used as an ETL to load data
- Add preprocess step in the model for 
- There's (always) a package for that!
- To insert code, use code snippets like below. If not evaluated as code, use {r eval=FALSE}. 
- For Stephs guides, go to lockedata.co.uk/firebrandmodule2

```{r}
install.packages('tidyverse')

```

# Multiple data sources
```{r}
library(tidyverse)

# Change "data" to where your files are.
# Remove the col_types bit if your columns
# are fairly type safe.
list.files("data", full.names = TRUE) %>%
  map_df(read_csv, col_types = cols("c")) %>% 
  bind_rows() %>% 
  nrow()
```
# Use caret package for reproducing transformation
For reproducing and operational purposes, the caret package is good. In the example below, all data that we input in the model gets transformed by "scale" (divides by the standard deviation), "center" (subtracts the mean of the predictor's data (again from the data in x) from the predictor values) and "pca" (use pca on the data). 

```{r}
install.packages(caret)
library(caret)

training_data    <- mtcars[1:20,-2]
training_outcome <- mtcars[1:20,2]
test_data        <- mtcars[21:32,-2]
test_outcome     <- mtcars[21:32,2]

model <- train(training_data, training_outcome,
              method="lm",
              preProcess = c("scale","center","pca"))
model
predictions <- predict(model, test_data)
predictions
```
#Plot data with ggplot

```{r}
install.packages("datasauRus")
library(ggplot2)
library(datasauRus)

myPlot<- ggplot(datasaurus_dozen, aes(x,y)) +
           geom_point()
myPlot
```


```{r}
library("tidyverse")
tidyverse_packages()
```

# Vectors
1d objects

```{r}
colours <- c("blue","red","grey")
colours2 = c("purple","green")

1:10
nums <- 1:10

about <- c(firstname="Eva", lastname="wilhelmsson")
about

```


Extracting values uses a grid system
- Recieving the first value in the vector, use for example colours[1] for extracting the first value
- All except the first value in the vector, use colours[-1]
- runif() 

```{r}
colours[1]
colours[-1]
colours[c(1,3)]
colours[1:2]
colours[-c(1,3)]
colours[runif(n=2, min=1, max=3)] #random no from the uniform distr
multcolours = c(colours,colours2)
multcolours
multcolours[runif(n=2, min=1, max=5)]
rm(colors,colors2)
```


This is an example of using python inside R studio. 

```{python}
import io
print("hello")
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
