---
title: "Visualizing data"
output: html_notebook
---

32.1.1 Exercise
Plot the iris dataset with Sepal.Width and Sepal.Length as the axes;
Try out some of the aesthetics used to see if there’s any clusters in the data;
Try out some of the aesthetics used to see if there’s any overplotting issues.


```{r}
ggplot(data = iris, aes(x=Sepal.Length, y=Sepal.Width))+geom_point()

ggplot(data = iris, aes(Sepal.Length, Sepal.Width, colour=Species))+geom_point()

ggplot(data = iris, aes(Sepal.Length, Sepal.Width, colour=Species))+geom_point(alpha=0.7)

```
32.2.1 Exercises
Produce a bar chart showing the max y value for each dataset in the datasaurus_dozen.

```{r}
ggplot(data = datasaurus_dozen, aes(x=rownames(datasaurus_dozen, y=)))+geom_bar()

ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg, colour = as.factor(am))) + 
    geom_col()

datasaurus_dozen %>%
    group_by(dataset) %>%
    summarise(y=max(y)) %>%
    ggplot(aes(x=dataset, y=y)) 
    + geom_col() 
    + theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```
#Density

```{r}
ggplot(data = mtcars, aes(x = wt)) + geom_histogram()
ggplot(data = mtcars, aes(x = wt)) + geom_density()

ggplot(data = mtcars, aes(x = as.factor(am), y = wt)) + geom_violin()

ggplot(data = mtcars, aes(x = wt, mpg)) + geom_bin2d()
```

33.0.1 Exercises
Build a density chart of your dd_sample using geom_density(), use the sorts of aesthetics values we’ve already used to make this a density by dataset.

```{r}
ggplot(data = dd_sample, aes())
```
#Facetting
Divide plots by grouping in the data by using facet_wrap. In the facet_wrap, there are multiple inputs in the facet_wrap. 

```{r}
ggplot(data = datasaurus_dozen, aes(x = x, y = y)) + geom_point() + facet_wrap(~dataset)
```
34.3 Exercises
Produce a facetted scatter plot for simpsons_paradox split by dataset.
```{r}
ggplot(data = simpsons_paradox, aes(x = x, y = y)) + geom_point() + facet_wrap(~dataset)

```
#Themes
Use some of the theme_xx or using 

```{r}
ggplot(data = mtcars, aes(x = cyl, y = wt, group = am, fill = am)) + geom_col() + 
    theme_dark()

ggplot(data = mtcars, aes(x = cyl, y = wt, group = am, fill = am)) + geom_col() + 
    ggthemes::theme_fivethirtyeight()

```

