---
title: "R Basics Firebrand"
output:
  html_notebook: default
---

# About R
- R came from S in the 70s
- R works in memory
    + Microsoft has out-of-memory processing capabilities
    + There's some other potential potential options
- R is querky
- R is case sensitive
- R can be used as an ETL to load data
- Add preprocess step in the model for 
- There's (always) a package for that!
- To insert code, use code snippets like below. If not evaluated as code, use {r eval=FALSE}. 
- For Stephs guides, go to lockedata.co.uk/firebrandmodule2

```{r}
install.packages("tidyverse")

```

# Multiple data sources

```{r}
library(tidyverse)

# Change "data" to where your files are.
# Remove the col_types bit if your columns
# are fairly type safe.
list.files("data", full.names = TRUE) %>%
  map_df(read_csv, col_types = cols("c")) %>% 
  bind_rows() %>% 
  nrow()
```
# Use caret package for reproducing transformation
For reproducing and operational purposes, the caret package is good. In the example below, all data that we input in the model gets transformed by "scale" (divides by the standard deviation), "center" (subtracts the mean of the predictor's data (again from the data in x) from the predictor values) and "pca" (use pca on the data). 

```{r}
install.packages(caret)
library(caret)

training_data    <- mtcars[1:20,-2]
training_outcome <- mtcars[1:20,2]
test_data        <- mtcars[21:32,-2]
test_outcome     <- mtcars[21:32,2]

model <- train(training_data, training_outcome,
              method="lm",
              preProcess = c("scale","center","pca"))
model
predictions <- predict(model, test_data)
predictions
```
#Plot data with ggplot

```{r}
rm()
install.packages("datasauRus")
library(ggplot2)
library(datasauRus)

myPlot<- ggplot(datasaurus_dozen, aes(x,y)) +
           geom_point()
myPlot

ggplot(data = simpsons_paradox, aes(x = x, y = y, group = dataset),colour = dataset) + geom_line() + geom_col()

ggplot(data = simpsons_paradox, aes(x = x, y = y, group = dataset, colour = dataset)) + 
    geom_jitter(width=5)

ggplot(data = simpsons_paradox, aes(x = x, group = dataset, colour = dataset)) + 
    geom_bar()

ggplot(data = mtcars, aes(x = rownames(mtcars), y = mpg, colour = as.factor(am))) + 
    geom_col()+theme(axis.text.x = element_text(angle = 90)) 
```

Use multiple charts

Some info of simpsons paradox: The Paradox is named after the statistician who discovered something unusual in the 1960s. It’s a great example of how stats can be wrong. The paradox is that averages can be silly and misleading. Sometimes they can be just plain baffling. (from http://www.statisticshowto.com/what-is-simpsons-paradox/)

By using %+%, we change the dataset in a model to another dataset. In the below example we change out dataset to simpson_paradox dataset in the model myPlot. 

```{r}
myPlot + facet_wrap(~dataset)

myPlot %+% simpsons_paradox
```


```{r}
library("tidyverse")
tidyverse_packages()
```


# Change from R to Python in the notebook 
This is an example of using python inside R studio. 

```{python}
import io
print("hello")
```
# Data types and basic operations

Basic sequence	:	1:3 = 1, 2, 3
Integer division	%/%	9 %/% 4 = 2
Modulus	%%	9 %% 4 = 1

If you’re comparing decimal values that result from calculations it is better to use the all.equal() function. all.equal() adds a tolerance to the comparison which means the very subtle imprecision is ignored. The default tolerance is 1.5 × 10^8. 


```{r}
1:3
9 %/% 4
9 %% 4
```
# LOGIC & Common operations
The ampersand (&) allows us to combine two checks to do an AND check, which is “are both things true?”.

The pipe, or bar (|) allows us to do an OR check, which is “are either of these things true?”.

The exclamation point (!) allows us to a perform a NOT check, by negating or swapping a check’s result. This allows you say things like “is this check true and that check not true?”.

```{r}
(2 < 3) & (4 == 4)
(2 < 3) & (4 != 4)
(2 < 3) | (4 == 4)
(2 < 3) | (4 != 4)

TRUE & TRUE
TRUE & !FALSE
!FALSE & !FALSE
(2 < 3) & (4 == 4)
(2 < 3) & !(4 != 4)



```
What is the result of pi^2?
Is pi greater than 3?
Construct a statement to check if 5 is both greater than 3 and less than or equal to 6
What are the results if you check to see if a sequence of 1 to 5 is less than or equal to 3?

```{r}
pi^2
pi>3

( 5>3 ) & (5<=6)

(1:5)<=3

```

# Vectors
1d objects

```{r}
colours <- c("blue","red","grey")
colours2 = c("purple","green")

1:10
nums <- 1:10

about <- c(firstname="Eva", lastname="wilhelmsson")
about

```


#Extracting values uses a grid system
- Recieving the first value in the vector, use for example colours[1] for extracting the first value
- All except the first value in the vector, use colours[-1]
- runif() 

```{r}
colours[1]
colours[-1]
colours[c(1,3)]
colours[1:2]
colours[-c(1,3)]
colours[runif(n=2, min=1, max=3)] #random no from the uniform distr
multcolours = c(colours,colours2)
multcolours
multcolours[runif(n=2, min=1, max=5)]

```
#Vector handles 
Element-wise or pair-wise calculation methodology.

```{r}
vecA <- 1:2
vecA
vecB <- 2:3
vecB
vecA*vecB
```

# Data frames

```{r}
baddf <- data.frame(a = 1:2,
                   b = c("blue","red"),
                   row.names = c("First","Second"))

gooddf <- data.frame(a = 1:2,
                   b = c("blue","red"),
                   ID = c("First","Second"))
View(iris)
class(iris)

mydf <- data.frame(a = 1:2, 
                   b = c("blue","red"), 
                   row.names = c("First","Second"))

names(mydf)
colnames(mydf)
rownames(mydf)

mylist <- list(a = 1:3, LETTERS)
mylist
```
#Working with functions
Examine the function by writing it without the brackets
```{r}
Sys.Date

```
